Case study: Analysis of smart meter data

### 2019-08-26 smart_meter.Rmd
only 9 months analysis included
adding analysis of weekends across weeks of the months

# Case study: Analysis on smart meter data {#smartmeter}

@@

Let us compare the energy consumption distribution for two households from the customer trials. For this analysis, we only take the data for the year 2013. Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.

@@

```{r cust1_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the first household. 10th, 25th, 50th, 75th and 90th percentile are shown. For example, 90th percentile might denote behavior which are extreme and 50th percentile corresponds to a more typical behavior. Evening hours generally exhibits a peak. Some extreme behavior is exhibited in the evening hours of May and June, however, overall energy comsumption does not seem to vary much across months. Median and 75th percentile is less than 0.25 KwH across most months."}
library(ggplot2)
data_sm1 %>% 
  granplot("month_year", "hour_day", 
           response = "general_supply_kwh", 
           plot_type = "quantile", 
           quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household. 10th, 25th, 50th, 75th and 90th percentile are shown. Distinct evening peaks for few hours across all months. The median comsumption also varies across months. This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January. This is evident from the 75th and 90th percentile of consumption."}
data_sm3 %>% 
  create_gran("month_year") %>% 
  filter(month_year != "Mar") %>% 
    granplot("month_year", "hour_day", "general_supply_kwh",
             plot_type = "quantile", 
             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + 
  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

@@

data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3)) +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```
<!-- # ``````{r cust2_hour-day_day-week} -->
<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
<!-- # ``` -->

```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household. Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent. Hours in between these have a long right tail which are even higher for weekends."}
data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3)) +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

@@

@@

 - First household typically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June). They are strictly 9 to 5 professionals, hence their energy consumption during day time is pretty consistent and only peaks up during a few hours in the evening. They do all energy-intensive chores (e.g. laundry, vacuum cleaning) mostly on the weekends.

 - Second household typically use air conditioners in summer months and heaters in winter months. This might be a family with children with stay-at-home parents/ work-from-home professionals with extremely varying energy consumption peaking up from noon. Their weekday behavior is not distinctly different from weekends. Also, they are more likely to distribute their energy-intensive chores throughout the week and not leave all them for the weekends.

@@

data_sm1 %>% granplot("month_year", 
                      "day_week", 
                      response = "general_supply_kwh", plot_type = "lv") +    ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2") + scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

>>>>

@@

Let us compare the energy consumption distribution for two households from the customer trials. For this analysis, we only take the data for the year 2013. The first household has data from April and the second household has data from March. For the purpose of comparison, data for both households have been observed from April. Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.

@@

```{r cust1_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the first household. 10th, 25th, 50th, 75th and 90th percentile are shown. For example, 90th percentile might denote behavior which are extreme and 50th percentile corresponds to a more typical behavior. Evening hours generally exhibits a peak. Some extreme behavior is exhibited in the evening hours of May and June, however, overall energy comsumption does not seem to vary much across months. Median and 75th percentile is less than 0.25 KwH across most months."}
library(ggplot2)
data_sm1 %>% 
  granplot("month_year", "hour_day", 
           response = "general_supply_kwh", 
           plot_type = "quantile", 
           quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  #ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +ggtitle("") +
  scale_x_discrete(breaks = seq(0,23,2))
```

```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household. 10th, 25th, 50th, 75th and 90th percentile are shown. Distinct evening peaks for few hours across all months. The median comsumption also varies across months. This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January. This is evident from the 75th and 90th percentile of consumption."}
data_sm3 %>% 
  create_gran("month_year") %>% 
  filter(month_year != "Mar") %>% 
    granplot("month_year", "hour_day", "general_supply_kwh",
             plot_type = "quantile", 
             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("") +
  #ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

@@

data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("") +
  #ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") 
  scale_x_discrete(breaks = seq(0,23,3)) +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```
<!-- # ``````{r cust2_hour-day_day-week} -->
<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
<!-- # ``` -->

```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household. Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent. Hours in between these have a long right tail which are even higher for weekends."}
data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +
  #ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") +
  scale_x_discrete(breaks = seq(0,23,3)) +
  ggtitle("") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

@@

data_sm1 %>% granplot("month_year", 
                      "day_week", 
                      response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +  #ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
  #ggtitle("LV plot of energy consumption across days of week by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

-  Across week of the month and weekends

Another aspect which might of interest is how these households distribute their monthly chores between Saturday and Sunday across weeks of the month. "Monthly" chores might include washing mattress, comforters, bed and pillow covers, vacuum cloth furniture or cleaning dishwasher or oven. Also some households might visit friends and family or have them come over on weekends at least once a month. Figures \ref{fig:cust1_wm_wndwd} and \ref{fig:cust3_wm_wndwd} shows that the letter value plot of energy consumption for Saturday and Sunday across weeks of the months.

Figure \ref{fig:cust1_wm_wndwd} shows that median consumption of energy is higher on Sundays. Also Sundays have longer right-tails than Saturdays across all weeks of the month implying that Sunday is a little more active day for the household. And the forth weekend of the month, either they take it easy on themselves or stay out of their house for trips/visit friends.

Figure \ref{fig:cust3_wm_wndwd} shows that household 2 has all-absorbing Sundays. They are either busy with weekly chores or have the entire family stay at home every Sunday. The first Saturday of the month looks pretty consistent compared to any other weeks of the month, implying either they are out for work/trips/visiting family. Also, 4th Sunday of the month has a long right tail implying they choose the last week of the month for doing their "monthly" chores.

```{r cust1_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 1. The 4th week of the month sees low energy consumption for both Saturdays and Sundays.Typically Sundays are heavier on energy than Saturdays."}

data_sm1 %>% 
   create_gran("day_week") %>% 
  filter(day_week %in%c("Sat", "Sun")) %>% 
  granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
  #ggtitle("Boxplot plot of energy consumption across days of week by month of years for Household 2") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

```{r cust3_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 2. The forth  Sunday has high energy consumption and long right tails implying they choose the last Sunday of the month either to do monthly chores or have friends come over. The first Saturday of the week seems quite and consistent for this household."}

data_sm3 %>%
  create_gran("day_week") %>% 
  filter(day_week %in%c("Sat", "Sun")) %>% 
  granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") + ggtitle("") +
  #ggtitle("Boxplot plot of energy consumption across days of week by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

@@

 - First household typically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June). They are strictly 9 to 5 professionals, hence their energy consumption during day time is pretty consistent and only peaks up during a few hours in the evening. They do all energy-intensive chores (e.g. laundry, vacuum cleaning) mostly on the weekends.
Also, part of the forth weekend of the month is spent out of the house visiting friends/family or going out for leisure trips.

 - Second household typically use air conditioners in summer months and heaters in winter months. This might be a family with children with stay-at-home parents/ work-from-home professionals with extremely varying energy consumption peaking up from noon. Their weekday behavior is not distinctly different from weekends. Also, they are more likely to distribute their energy-intensive chores throughout the week and not leave all them for the weekends. Sundays are still pretty heavy on energy for this household when all members stay at home. Also, they sometimes spend the last Sunday of the month doing "monthly" chores or have people come over.

@@
